---
title: 'Outlier Detection: Isolation Forest (Penguins)'
author: "Darian-Florian Voda"
date: "2022-12-30"
output:
  html_document:
    df_print: paged
  pdf_document: default
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

## Loading packages
```{r}
library(caret)
library(dplyr)
library(mltools)
library(rpart)
library(isotree)
library(plotly)
```

## Exercise

Let's check if there are anomalies in the dataset *"PenguinsWithoutMissingValues.csv"*. We want to use Isolation Forests for it since they have some advantages compared with other anomaly detection techniques.

- import the data PenguinsWithoutMissingValues

```{r}
# import the data PenguinsWithoutMissingValues
setwd("C:/Users/Dari-Laptop/Desktop/FH Karnten - Master - AppDs/StatisticsAppDSLaptop")
data = read.csv("PenguinsWithoutMissingValues.csv")
```
- encode categorical data

```{r}
# encode categorical data
data$Species = factor(data$Species)
data$Island = factor(data$Island)
data$Gender = factor(data$Gender)
```

- predict the average depth of the isolation forest (using the R library isotree)

```{r}
# predict the average depth

model <- isolation.forest(data, ndim=1, ntrees=10, nthreads=1)
scores <- predict(model, data, type="avg_depth")
par(mar = c(4,5,3,2))
plot(data$CulmenLength.mm., scores, type="p", col="darkred",
     main="Average isolation depth\nfor CulmenLength(mm)",
     xlab="value", ylab="Average isolation depth")


plot(data$CulmenDepth.mm., scores, type="p", col="darkred",
     main="Average isolation depth\nfor CulmenDepth(mm)",
     xlab="value", ylab="Average isolation depth")

plot(data$FlipperLength.mm., scores, type="p", col="darkred",
     main="Average isolation depth\nfor FlipperLength(mm)",
     xlab="value", ylab="Average isolation depth")

plot(data$BodyMass.g., scores, type="p", col="darkred",
     main="Average isolation depth\nfor BodyMass(g)",
     xlab="value", ylab="Average isolation depth")
```

- calculate the anomaly score (see slides or original paper)

```{r}
data$score <- predict(model, newdata = data)

data$score
```

- use ggplotly scatterplot to visualize the anomaly scores and show the IndividualIDs in the hoover text

```{r}
# use ggplotly scatterplot to visualize the anomaly scores 
# and show the IndividualIDs in the hoover text

p = ggplot(data, 
           aes(x=score, y = 1:nrow(data), color=Species, 
               text = paste("IndividualID :", IndividualID))) + 
  geom_point() + 
  labs(title="Anomaly scores",
                      x ="Score", y = "Number")

p

ggplotly(p)

```

## Conclusions

- Yes, there are anomalies regarding to our anomaly scores
- I have computed the anomaly score using the predict function and not from scratch, even though several trials occured using the anomaly formula from the Isolation Forests article
- ID number N81A1 seems to be an anomaly from its species
- Same applies for ID N88A1, N72A1 and many others as we can see from our plot

