---
title: "Homework 7"
author: "Darian-Florian Voda"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Loading packages
```{r}
library(dplyr)
library(ggplot2)
library(car)
library(tidyverse)
library(ggpubr)
library(rstatix)
```

## Exercise 85: Outcome of lung ventilation

- Use the dataset „Discriminant-pulmonary.txt“
- Contains diverse parameters on patients receiving lung ventilation
  - Age
  - Oxygen concentration in blood
  - Body size
  - Aggressiveness of the ventilation
  - Ventilation time
- Not all patients survived. Can you predict, based on these parameters, whether a patient will survive or not?
- Also try standardization of variables!

Load packages
```{r}
library(ggplot2)
library(klaR)
library(ggord)
library(psych)
library(MASS)
library(devtools)
library(tidyverse)
library(caret)
library(mosaic)
library(broom)
```

For this exercise, we will try to predict **using** scaled variables and **without** scaled variables.

Let's start without scaled variables
```{r}
# Without standardization
data<-read.delim("C:/Users/daria/OneDrive/Desktop/Master - AppDS/Statistics/Datasets-20221007/Discriminant-pulmonary.txt",
                stringsAsFactors=F)
# Look at the data
head(data)

# Get type of the data
str(data)

pairs.panels(data[2:6],
             gap = 0,
             bg = c("red", "green")[data$outcome],
             pch = 21)

# 70% for training, 30% for testing
set.seed(123)
ind <- sample(2, nrow(data),
                replace = TRUE,
                prob = c(0.7, 0.3))
training <- data[ind==1,]
testing <- data[ind==2,]

linear <- lda(outcome~., training)
linear

plot(linear)

paste("Prior probabilities of survival: 50%/50% rate of survival")

# prediction
p <- predict(linear, training)
names(p)

# Predicted classes
head(p$class, 6)

# Predicted probabilities
head(p$posterior, 6)

# Linear discriminants
head(p$x, 3)

# Stacked histogram for LD1
p <- predict(linear, training)
ldahist(data = p$x[,1], g = training$outcome)

lda.data <- cbind(training, predict(linear)$x)
ggplot(lda.data, aes(LD1, age)) +
  geom_point(aes(color = outcome)) +
  labs(title="Outcome of patients with pulmonary problems")

# plot the classes using density plot
ggplot(data = lda.data)+
  geom_density(aes(age, fill = outcome), alpha = 0.2)+
  scale_fill_discrete(name = "Outcome of patients", labels = c("Dead", "Survived"))

partimat(factor(outcome)~., data = training,
         method = "lda")

# prediction accuracy of training set
p1 <- predict(linear, training)$class
tab <- table(Predicted = p1,
               Actual = training$outcome)
tab
sum(diag(tab))/sum(tab)


# prediction accuracy of test set
p2 <- predict(linear, testing)$class
tab1 <- table(Predicted = p2,
                Actual = testing$outcome)
tab1
sum(diag(tab1))/sum(tab1)

# QDA
quadratic <- qda(outcome~., data = training)
quadratic


predquad <- predict(quadratic, training)
names(predquad)

# prediction accuracy of training set
pq1 <- predict(quadratic, training)$class
tab <- table(Predicted = pq1,
               Actual = training$outcome)
tab
sum(diag(tab))/sum(tab)

# prediction accuracy of test set
pq2 <- predict(quadratic, testing)$class
tab1 <- table(Predicted = pq2,
                Actual = testing$outcome)
tab1
sum(diag(tab1))/sum(tab1)

partimat(factor(outcome)~., data = training,
         method = "qda")
```
From this we can conclude that:

- Prior probabilities of groups are
  - 50% chance to die
  - 50% chance to survive
- LD1 = -0.0085 * age - 2.164 * oxygen.conc + 0.032 * body.size - 0.054 * ventilation.aggressiveness - 0.089 * ventilation.time
- The stacked histogram seems to show us that both groups are overlapping
- However, we can predict with a 68% accuracy on the training set and 67% on the test set, which is a good accuracy despite the overlapping
- Using QDA we get 74% accuracy on the training set and 67% on the test set, which is also an improvement on our learning algorithm.

Now let's try with **standardized** variables:
```{r}
### data with standardization
data2<-read.delim("C:/Users/daria/OneDrive/Desktop/Master - AppDS/Statistics/Datasets-20221007/Discriminant-pulmonary.txt",
                 stringsAsFactors=F)

data2_standardized = data2 %>% mutate_at(c('oxygen.conc', 'body.size', 'ventilation.aggressiveness', 'ventilation.time', 'age'),
                         ~(scale(.) %>% as.vector))

head(data2_standardized)


pairs.panels(data2_standardized[2:6],
             gap = 0,
             bg = c("red", "green")[data2_standardized$outcome],
             pch = 21)

# 70% for training, 30% for testing
set.seed(123)
ind <- sample(2, nrow(data2_standardized),
              replace = TRUE,
              prob = c(0.7, 0.3))
training <- data2_standardized[ind==1,]
testing <- data2_standardized[ind==2,]

linear <- lda(outcome~., training)
linear

plot(linear)

paste("Prior probabilities of survival: 50%/50% rate of survival")

# prediction
p <- predict(linear, training)
names(p)

# Predicted classes
head(p$class, 6)

# Predicted probabilities
head(p$posterior, 6)

# Linear discriminants
head(p$x, 3)

# Stacked histogram for LD1
p <- predict(linear, training)
ldahist(data = p$x[,1], g = training$outcome)

lda.data <- cbind(training, predict(linear)$x)
ggplot(lda.data, aes(LD1, age)) +
  geom_point(aes(color = outcome)) +
  labs(title="Outcome of patients with pulmonary problems")

# plot the classes using density plot
ggplot(data = lda.data)+
  geom_density(aes(age, fill = outcome), alpha = 0.2)+
  scale_fill_discrete(name = "Outcome of patients", labels = c("Dead", "Survived"))

partimat(factor(outcome)~., data = training,
         method = "lda")

# prediction accuracy of training set
p1 <- predict(linear, training)$class
tab <- table(Predicted = p1,
             Actual = training$outcome)
tab
sum(diag(tab))/sum(tab)


# prediction accuracy of test set
p2 <- predict(linear, testing)$class
tab1 <- table(Predicted = p2,
              Actual = testing$outcome)
tab1
sum(diag(tab1))/sum(tab1)

# QDA
quadratic <- qda(outcome~., data = training)
quadratic


predquad <- predict(quadratic, training)
names(predquad)

# prediction accuracy of training set
pq1 <- predict(quadratic, training)$class
tab <- table(Predicted = pq1,
             Actual = training$outcome)
tab
sum(diag(tab))/sum(tab)

# prediction accuracy of test set
pq2 <- predict(quadratic, testing)$class
tab1 <- table(Predicted = pq2,
              Actual = testing$outcome)
tab1
sum(diag(tab1))/sum(tab1)

partimat(factor(outcome)~., data = training,
         method = "qda")
```

We conclude from this standardization that:

- Both not standardized and standardized methods show the same accuracy on predictibility
- Standardization is better for understanding the importance of our data

## Exercise 86

- In bird species where males and females are coloured the same, it is usually difficult to determine sex on the basis of external characteristics such as size or behaviour.
- Therefore, either an endoscopic examination of the internal reproductive organs or a blood sample must be taken.
- Both procedures put the animals under a lot of stress, so they usually take place under general anaesthesia, which can also affect the birds’ health.
- In addition, analysing the blood for hormone status or certain genetic traits is costly and requires specialist staff.
- The aim of this study is to determine the sex of the birds based on the characteristics of wing length, beak length, head length, foot length and weight.
- The aim of this study is to establish a discriminant function for sex determination using data from 245 birds, most of which have known sex.
- The following research questions need to be answered:
  - What is the discriminant function?
  - How many birds can be correctly assigned to their sex with this?
  - How many of the birds whose sex is unknown can still be assigned to a sex?

```{r}
#### Exercise 86 ####
data3<-read.delim("C:/Users/daria/OneDrive/Desktop/Master - AppDS/Statistics/Datasets-20221007/Discriminant-birds.txt",
                  stringsAsFactors=F)

# Look at the data
head(data3)

# Get type of the data
str(data3)

pairs.panels(data3[2:6],
             gap = 0,
             bg = c("red", "green")[data3$ï..sex],
             pch = 21)

# 70% for training, 30% for testing
set.seed(123)
ind <- sample(2, nrow(data3),
              replace = TRUE,
              prob = c(0.7, 0.3))
training <- data3[ind==1,]
testing <- data3[ind==2,]

linear <- lda(ï..sex~., training)
linear

paste("Prior probabilities of sex: 32% female, 15% male, 52% unknown")

# prediction
p <- predict(linear, training)
names(p)

# Predicted classes
head(p$class, 6)

# Predicted probabilities
head(p$posterior, 6)

# Linear discriminants
head(p$x, 3)

# Stacked histogram for LD1
p <- predict(linear, training)
ldahist(data = p$x[,1], g = training$ï..sex)

lda.data <- cbind(training, predict(linear)$x)
ggplot(lda.data, aes(LD1, LD2)) +
  geom_point(aes(color = ï..sex)) +
  labs(title="Sex of birds")

# Biplot
ggord(linear, training$ï..sex,
      ylim = c(-10, 10), xlim=c(-10,10))

partimat(factor(ï..sex)~., data = training,
         method = "lda")

# prediction accuracy of training set
p1 <- predict(linear, training)$class
tab <- table(Predicted = p1,
             Actual = training$ï..sex)
tab
sum(diag(tab))/sum(tab)


# prediction accuracy of test set
p2 <- predict(linear, testing)$class
tab1 <- table(Predicted = p2,
              Actual = testing$ï..sex)
tab1
sum(diag(tab1))/sum(tab1)

# QDA
quadratic <- qda(ï..sex~., data = training)
quadratic


predquad <- predict(quadratic, training)
names(predquad)

# prediction accuracy of training set
pq1 <- predict(quadratic, training)$class
tab <- table(Predicted = pq1,
             Actual = training$ï..sex)
tab
sum(diag(tab))/sum(tab)

# prediction accuracy of test set
pq2 <- predict(quadratic, testing)$class
tab1 <- table(Predicted = pq2,
              Actual = testing$ï..sex)
tab1
sum(diag(tab1))/sum(tab1)

partimat(factor(ï..sex)~., data = training,
         method = "qda")
```

Thus, we can conclude that:

- Prior probabilities of sex groups are: 
  - 32% female
  - 16% male
  - 52% unknown
- The discriminant functions are:
  - LD1 = 0.026 * wing - 0.260 * beak - 0.212 * head + 0.014 * food - 0.004 * weight
  - LD2 = 0.096 * wing + 0.180 * beak - 0.105 * head - 0.084 * food - 0.030 * weight
- There are quite no overlaps of the stacked histograms
- The scatter plot shows a better performance for QDA method
- There is 75% accuracy on the trained test and 64% accuracy on the test set, thus, a 64% chance to assign the sex correctly to a bird using LDA method
- There is 82% accuracy on the trained test and 63% accuracy on the test set, thus, a 63% chance to assign the sex correctly to a bird using QDA method
- There is a total of 123 birds whose sex is still unknown and still can be assigned. [Total sum of unknown column for both training and test set]